%% Package and Class "uiucthesis2014" for use with LaTeX2e.
\documentclass[edeposit,fullpage]{uiucthesis2018}

%\usepackage[utf8]{inputenc}


\usepackage[acronym,toc]{glossaries}
\makeglossaries
\include{acros}


\usepackage{xspace}
\usepackage{graphics}
\newcommand{\Cycamore}{\textsc{Cycamore}\xspace}
\newcommand{\Cyclus}{\textsc{Cyclus}\xspace}


\usepackage{placeins}
\usepackage{booktabs} % nice rules (thick lines) for tables
\usepackage{microtype} % improves typography for PDF

\usepackage[hyphens]{url}
\usepackage[hidelinks]{hyperref}
%\usepackage{subfig}
\usepackage{hhline}
\usepackage{amsmath}
\usepackage{color}
\usepackage{float}
\usepackage{multirow}
\usepackage{siunitx}
\usepackage{subcaption}
\usepackage{adjustbox}
\sisetup{ 
    input-decimal-markers = .,input-ignore = {,},table-number-alignment = right,
    group-separator={,}, group-four-digits = true
}
\usepackage{fourier}
\usepackage{booktabs}
\newcommand\tab[1][1cm]{\hspace*{#1}}

\usepackage{threeparttable, tablefootnote}

%tikzpicture fit to page width
\usepackage{environ}
\makeatletter
\newsavebox{\measure@tikzpicture}
\NewEnviron{scaletikzpicturetowidth}[1]{%
  \def\tikz@width{#1}%
  \def\tikzscale{1}\begin{lrbox}{\measure@tikzpicture}%
  \BODY
  \end{lrbox}
  \pgfmathparse{#1/\wd\measure@tikzpicture}%

  \edef\tikzscale{\pgfmathresult}%
  \BODY
}

\usepackage{tabularx}
\newcolumntype{b}{>{\hsize=1.0\hsize}X}
\newcolumntype{q}{>{\hsize=0.5\hsize}X}
\newcolumntype{R}{>{\raggedleft\arraybackslash\hsize=0.5\hsize}X}
\newcolumntype{z}{>{\hsize=0.75\hsize}X}
\newcolumntype{s}{>{\hsize=.5\hsize}X}
\newcolumntype{m}{>{\hsize=.75\hsize}X}

\usepackage{longtable}

\usepackage{cleveref}
\usepackage{datatool}
\usepackage[numbers]{natbib}
\usepackage{notoccite}

\usepackage{tikz}
\usetikzlibrary{positioning, arrows, decorations, shapes, shadings}

\usetikzlibrary{shapes.geometric,arrows}
\tikzstyle{process} = [rectangle, rounded corners, minimum width=2.5cm, minimum height=1cm,text centered, draw=black, fill=blue!30]

\tikzstyle{object} = [ellipse, rounded corners, minimum width=3cm, minimum height=1cm,text centered, draw=black, fill=green!30]
\tikzstyle{objectr} = [ellipse, rounded corners, minimum width=3cm, minimum height=1cm,text centered, draw=black, fill=red!30]

\tikzstyle{empty} =  [rectangle, rounded corners, minimum width=2.5cm, minimum height=0.7cm,text centered, draw=black, fill=white!30]
\tikzstyle{arrow} = [thick,->,>=stealth]

\newcommand\mainmatterWithoutReset
 {\edef\temppagenumber{\arabic{page}}%
  \mainmatter
  \setcounter{page}{\temppagenumber}%
 }

\title{Isotopic and Reactor Physics Characterization of a Gas-Cooled, Pebble-Bed Microreactor}
\author{Zo{\"e} Richter}
\department{Nuclear, Plasma, Radiological Engineering}
\schools{B.S., University of Illinois - Urbana Champaign, 2018}
\msthesis
\advisor{Madicken Munk}
\degreeyear{2021}
\committee{Dr. Madicken Munk \\ Dr. Tomasz Kozlowski}

% Template created by Charles Bae, 2018

\begin{document}
\maketitle

\frontmatter
%% Create an abstract that can also be used for the ProQuest abstract.
%% Note that ProQuest truncates their abstracts at 350 words.
\begin{abstract}
Pebble-bed \acrfull{htgr} designs present a unique modeling challenge.  Pebble-beds can have a variety of pebble compositions due to varying levels of burnup.  In addition, the pebbles are mobile in the core --- entering from the top and exiting through the bottom --- and may fall in a haphazard arrangement.  This work introduces a 20 MWth pebble-bed HTGR reactor design (which will be referred to as Sangamon20) and investigates not only the neutronics of the base model, but the changes to core neutronics after making modifications to the simulation.  These modifications include: heterogenizing versus homogenizing the pebble center, imposing a universal symmetry assumption, and changing the arrangement of pebble fuel compositions, all using Serpent and Python.  All of this is in support of the ultimate goal of this project --- to establish a baseline source term and to determine what simplifications, if any, can be made in the model without sacrificing accuracy in the results.  This is crucial for any future licensing effort, safety analysis, or accident analysis involving pebble-bed reactors.  This model of Sangamon20 uses a random dispersal of seven different pebble compositions, each corresponding to a different burnup level.  The heterogenization tests compare $k_{eff}$, thermal and fast flux profiles, and the neutron lethargy-adjusted energy spectra in the core, reflector, coolant, a random fresh, and a random discharge-burnup pebble.  Shuffling and symmetry tests monitor changes to $k_{eff}$ and the outward neutron current at the outer reflector boundary; the latter because it can be used to find the anticipated neutron flux the \acrfull{rpv} would experience.  This informs the level of radiation damage one could expect the \acrshort{rpv} to experience each year - which is useful from a design and safety perspective.  Neither the symmetry test nor the shuffling test caused a major difference in either the $k_{eff}$ or the outer-bounds outward neutron current.  This would suggest that there is no need to simulate all possible pebble placements to characterize a reactor.  However, for the heterogenization tests, $k_{eff}$ differed by over 4.0\%, and the pebble spectra at certain higher energies disagreed by a factor of 2-4.  A complete fuel isotopic composition is accessible at \cite{richter_isotopic_2021}, and this thesis discusses select isotopic inventories.  These can be used to inform source term determination or spent fuel compositions.

\end{abstract}

\chapter*{Acknowledgments}
\input{acks}

%% The thesis format requires the Table of Contents to come
%% before any other major sections, all of these sections after
%% the Table of Contents must be listed therein (i.e., use \chapter,
%% not \chapter*).  Common sections to have between the Table of
%% Contents and the main text are:
%%
%% List of Tables
%% List of Figures
%% List Symbols and/or Abbreviations
%% etc.

\tableofcontents
\listoftables
\listoffigures
\addcontentsline{toc}{chapter}{List of Abbreviations}

%% Create a List of Abbreviations. The left column
%% is 1 inch wide and left-justified
\chapter*{List of Abbreviations}
\input{acros-list}
%\printglossary
%\printglossary[type=\acronymtype, title=List of Acronyms, toctitle=List of Acronyms]
%% Create a List of Symbols. The left column
%% is 0.7 inch wide and centered

\pagebreak
\mainmatter

\chapter{Introduction}
\input{introduction}

\chapter{Literature Review}
\input{litreview}

\chapter{Methodology}
\label{sec:methods}
\input{methods}

\chapter{Results}
\input{results}

\chapter{Conclusion}
\input{conclusion}

\backmatter
\nocite{*}
\bibliographystyle{IEEEtran}
\bibliography{bibliography}

\mainmatterWithoutReset
\appendix
\chapter{}
\input{appendix}

\chapter{}
\input{appendixB}


\end{document}
\endinput
%%
%% End of file `thesis-ex.tex'.
